---
title: "simpleRegress"
author: "Lele Nicholson"
date: "Fall 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
Seeing the predicted weight of the bear compared to the neck has a standered error of 43.91 and the multiple R squared is 0.8725. The neck's error is 1.069 while its intercepted or the other error of 22.779.

## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted weight of the bear compared to the chest. The Adjusted R-squared is 0.9262 and the Residual standard error is 33.08.

```{r}
simp <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted weight of the bear compared to the length. The Adjusted R-squared 0.7423  and the Residual standard error 0.7472

```{r}
simp <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted weight of the bear compared to the Headwth. The Adjusted R-squared 0.6064 and the Residual standard error 0.6138. 

## CONCLUSION AND RANKING

1. the predicted weight and chest. The AR-squared: 0.9262 and the Rse: 33.08.
2. the predicted weight and length. The AR-squared: 0.7423 and the Rse: 0.7472.
3. the predicted weight and HEADWTH The AR-squared: 0.6064 and the Rse: 0.6138.
